---
title: "ESM 244 Lab 2"
author: "Allison Horst"
date: "December 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(janitor)
library(MASS)
library(ggrepel)
library(RColorBrewer)
library(factoextra)
library(ggbiplot)
library(shinythemes)
```
 if u want to override (set) a function manually throut an entire workspace you can assign the function name explicitly
 
```{r}
select <- dplyr::select
```
 
 
 #1. ordinal logistic regression for political party affiliation by ethnicity age and sex

 

```{r, message = FALSE}

pol_party <- read_csv("pol_party.csv") %>% 
  clean_names() # janitor function for snake case

pol_df <- pol_party %>% 
  select(birth_year, gender, ethnicity, political_party) %>% 
  filter(birth_year > 1900, political_party != "NA", political_party <= 7) %>% 
  mutate(age = 2016 - birth_year)

# For political party: 1 = very liberal, 4 = moderate, 7 = very conservative

# Coerce to factors (including ordered factor for the ranking of political party)

pol_df$political_party <- factor(pol_df$political_party, ordered = TRUE, levels = c("1","2","3","4","5","6","7")) # Check class and levels to ensure

# Then update the gender and ethnicity variables to factors (unordered)

pol_df$ethnicity <- factor(pol_df$ethnicity)
pol_df$gender <- factor(pol_df$gender) 

# Then we can run ordinal logistic regression, but we should do some visualization first.

```

- Exploratory visualization

```{r, message = FALSE}

counts_eth <- pol_df %>% 
  group_by(ethnicity, political_party) %>% 
  tally()

counts_g <- pol_df %>% 
  group_by(gender, political_party) %>% 
  tally()

# Exploratory graphs

ggplot(pol_df, aes(x = age, y = political_party)) +
  geom_point(alpha = 0.5, aes(color = ethnicity)) # no notable effect of age

ggplot(pol_df, aes(x = political_party)) +
  geom_histogram(aes(fill = ethnicity), stat = "count") +
  scale_fill_brewer(palette = "YlGnBu") +
  theme_light() +
  theme(legend.position = "NA") +
  facet_wrap(~ethnicity, scales = "free")
```
n is too small for groups
```{r}
pol_model <- polr(political_party ~ age + gender + ethnicity, data= pol_df)

summary(pol_model)
# hard to find p value cause t value so its not even good we dont care about it
#logodds coeff of 0= no effect so close to zero is sooooooo teeny
#male vs female: males more likely to be conservative (log odds of 0.31)
#if n is so small for a group the log odds will have no meaning like for native american group

exp(coef(pol_model))
#1 = no change
#age so close to 1 so barely any effect
#baselevel= female and asian


#THIS IS NOT PUBLISHABLE DATA!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!! SO BIAS!!!! BAD BAD BAD SAMPLE SIZE IS TOO SMALL GO TO JAIL


#once weve decided on a model we can use it to make predictions

predictions <- predict(pol_model, type="probs")
#can change type to display diff things like odds or logodds 
df<- data.frame(pol_df, predictions)
df

#can look at a hypothetical individual ex white woman age 26
```

```{r}
wb_10 <- read_csv("wb_10.csv") %>% 
  select(-X1) %>% 
  column_to_rownames('code')

wb_pca <- prcomp(wb_10[3:8], scale=TRUE)

summary(wb_pca)

ggbiplot(wb_pca) +
  geom_text_repel(label = rownames(wb_10), size = 2, segment.size = 0.2) +
  theme_bw() 

ggbiplot(wb_pca, groups = wb_10$region) +
  geom_text_repel(label = rownames(wb_10), size = 2, segment.size = 0.2) +
  theme_bw() 
```

#factoextra package was built to help visualize outcomes of ordination methods
```{r}

# Scree plot!
fviz_screeplot(wb_pca)

# Show contributions/quality of representation for different variables on the PCs:

# Contributions of variables to PC1
fviz_contrib(wb_pca, choice = "var", axes = 1, top = 6)
# Contributions of variables to PC2
fviz_contrib(wb_pca, choice = "var", axes = 2, top = 6)

#biplot showing points and vectors

fviz_pca_biplot(wb_pca,
                col.var ="cos2",
                label = "var",
                repel= TRUE)
#cos2= higher = better represented in biplot from first variable component
```


DAS SHINY APP!!!!  brackets and parentheses are EVERYTHING!!!
```{r}

```


